---
title: "Apply R"
author: "Anatoly Dryga"
date: "2/9/2016"
output: pdf_document
---

# Prelim

What are data types:

```{r}
str("1")
str(1)
str(1.0)
str(1L)
```

why this is important?.. think about memory required to store value

If we have __multiple__ things:

* of the same type(can be stored/accessed efficiently)
hidden from us completely in R(errrr)

```{r}
v <- c(1,2,3)
str(v)
```

* of different types

```{r}
L <- list(1,"2",3L)
str(L)
```

# lapply, sapply

Looping: repeating something over and over.

lapply/sapply  - just loops hidden/written differently
(this statement can start a **flame** war).

Starting with vector and returning the vector:

```{r}
add100 <- function(x) {x + 100}
sapply(v, add100)
```
(in words applying function add100 for each element of v)

Or using anonymous(if you don't need a name) function:
```{r}
sapply(v, function(x) { x + 100})
```

which is really similar to(but we are not returning vector):
```{r}
for (element in v) print(add100(element))
```

And now lapply returns list:
```{r}
lapply(v, function(x) { x + 100})
```

that fails:

```{r, eval=FALSE}
lapply(L, function(x) { x + 100})
```
Why? For what data types '+' is defined?

# (Beautiful) Hacks

* a dataframe can be used as list of columns:

```{r}
library(lattice)
str(barley)
lapply(barley, function(x) length(unique(x)))
```

* every operation is a function call
(and we can skip arguments - what a useful feature for bug intoduction)

```{r}
"+"(2,3)
A<-matrix(1:9, 3,3)
B<-matrix(4:15, 4,3)
C<-matrix(8:10, 3,2)
MyList<-list(A,B,C)
MyList
lapply(MyList, "[", , 2)
```

* use indexes and variable from the environment:

```{r}
sapply(2:3, function(x) { sum(A[, x])})
```

Arguably cleaner version:

```{r}
sapply(2:3, function(x, m) { sum(m[, x])}, A)
```

Why cleaner? globals are pure evil(most of the time)

# When to chose lapply vs sapply?

Still to come...







